## Intro

This code compares travel time and distance from Simulation output with Google Maps API
To use this code you need to have a valid Google Maps API key.
copy and paste your API key in the 'api/google.key' to be used below
Please note Google Maps API is NOT a free service - be aware of the potential costs

```{r setup, include=FALSE}
library(mapsapi)
library(sf)
library(lubridate)
library(tidyverse)
apikey <- readLines("./data/api-key.txt")

```

```{r Reading input from simulation}


simOutputTrips <- read_delim(gzfile("./data/simOutputs/output_trips.csv"),
                      delim=";") %>% 
  # Removing trips with zero distance
  filter(traveled_distance>0) 

network <- read_sf("./data/edgesMelbourne.gpkg") 

# Adding time of the day and date to the trips
```

Getting the distance and time for sampled trips for Google API

```{r get dist and time}

modesMap=tibble(modes=c("pt","car","truck","walk","bicycle"),
           gMode=c("transit","driving","driving","walking", "bicycling"))

tripsWithDepTime <- simOutputTrips %>% 
  left_join(modesMap, by=c("longest_distance_mode"="modes")) %>% 
  filter(gMode == "driving") %>% 
  mutate(depDate = paste0(Sys.Date() + 1, " ", dep_time)) %>% 
  mutate(depDate = as.POSIXct(depDate,format="%Y-%m-%d %H:%M:%OS",
                              tz = Sys.timezone())) %>% 
  filter(!is.na(depDate))
  

```

Sampling for each valid mode (adjust n)

```{r sampling the trips}

# Add this part to distinguish start links in urban and rural areas
network <- network %>%
  dplyr::select(link=edgeID, urban) %>%
  st_drop_geometry()
tripsWithDepTime$link <- as.numeric(gsub(".*?([0-9]+).*", "\\1", tripsWithDepTime$start_link))
tripsWithDepTime <- left_join(tripsWithDepTime, network, by= "link")

# Take Samples
n <- 500

tripsSampled <- tripsWithDepTime %>% 
  group_by(urban) %>%
  slice_sample(n = n) %>% 
  ungroup()

origins <- tripsSampled %>% 
  dplyr::select(trip_id,start_x, start_y) %>% 
  mutate(GEOMETRY=paste0("POINT(",start_x," ",start_y,")")) %>%
  st_as_sf(wkt = "GEOMETRY", crs = 28355) %>% 
  as.data.frame() %>%
  st_sf() %>% 
  st_transform(4283)

destinations <- tripsSampled %>% 
  dplyr::select(trip_id,end_x, end_y) %>% 
  mutate(GEOMETRY=paste0("POINT(",end_x," ",end_y,")")) %>%
  st_as_sf(wkt = "GEOMETRY", crs = 28355) %>% 
  as.data.frame() %>%
  st_sf() %>% 
  st_transform(4283)

```


## Assigning the Google Time and Distance to the Sample

```{r, warning=FALSE, message=FALSE}
i=1
for(i in 1:nrow(tripsSampled)){
# for(i in 1:10){
  if(i%%10 == 0) print(paste0("trip number ",i," out of ", nrow(tripsSampled), " trips"))
  route=mapsapi::mp_matrix(origins = origins[i,"GEOMETRY"],
                           destinations = destinations[i,"GEOMETRY"],
                           mode = tripsSampled$gMode[i],
                           departure_time = tripsSampled$depDate[i],
                           traffic_model="best_guess",
                           key = apikey)
  tripsSampled[i,"time_google"] <- route %>% mp_get_matrix(value = c("duration_s")) %>% as.numeric()
  tripsSampled[i,"distance_google"] <- route %>% mp_get_matrix(value = c("distance_m")) %>% as.numeric()
}
```


```{r}
tripsProcessed <- tripsSampled %>% 
  mutate(trav_time=seconds(trav_time)) %>% 
  mutate(time_google=seconds(time_google)) %>% 
  dplyr::select(person, trip_number, depDate, mode=longest_distance_mode, urban,
                 time_simulation=trav_time, distance_simulation=traveled_distance, 
                time_google, distance_google) %>%
  mutate(`Travel Distance Difference`=(distance_simulation-distance_google)) %>%
  mutate(`Travel Time Difference`=(time_simulation-time_google)) %>% 
  mutate(`Travel Time Error`=100*(time_simulation-time_google)/time_google) %>% 
  mutate(`Travel Distance Error`=100*(distance_simulation-distance_google)/distance_google)

tripsProcessed <-  tripsProcessed %>% 
  mutate(depDate = as.character(depDate))

write_csv(tripsProcessed, "./output/tripsProcessed.csv")
```

## Comparing the time and distance

### Time

```{r}

# Plot Differences

tripsProcessed$Travel.Time.Difference <- str_sub(tripsProcessed$Travel.Time.Difference, end = -2)
tripsProcessed$Travel.Time.Difference <- as.numeric(tripsProcessed$Travel.Time.Difference)/60

tripsProcessed %>%  ggplot(aes(x= Travel.Time.Difference)) +
  geom_histogram(colour = "white", fill= "blue", bins = 10,
                 breaks = seq(-20,20, by=5)) +
  scale_x_continuous( breaks = seq(-20,20, by=5)) + 
  xlab("Travel Time Diferrence (min)") +
  ylab("Count")

ggsave("output/TimeDiffHis.png",
       width = 5, height = 5)


```


```{r}

# Plot Error 
tripsProcessed %>% 
  filter(Travel.Time.Error<1000) %>%
  mutate(`Travel Mode`=case_when(mode=="car" ~ "Driving",
                                 mode=="truck" ~ "Driving",
                                 TRUE ~ "NA")) %>% 
  ggplot(aes(x=`Travel Mode`, y=Travel.Time.Error, 
             fill= `Travel Mode`)) + 
  ggtitle(" All Trips")+
  geom_boxplot(alpha=0.8) +
  scale_fill_brewer(palette = "Greys") +
    guides(colour = guide_legend()) +
    theme(legend.position ="none") +  
  scale_y_continuous(breaks = seq(0, max(500, na.rm = TRUE), by = 100))  

ggsave("output/Error.png",
       width = 5, height = 5)

```


